---
title: "Inferential Analysis"
author:
  - "Eva Kormann"
  - "Thomas Klebel"
format: html
editor: visual
execute: 
  echo: true
  warning: false
---

```{r}
#| output: false
#| echo: false

library(here)
library(tidyverse)
library(broom)


# Get var overview including short titles
vars <- here("data", "processed", "var_overview.csv")
var_overview <- read_csv(vars)

# Get functions for analysis
functions <- here("R", "analysis.R")
source(functions)

# Get data
use_targets <- TRUE # Set to false if data is not received by targets pipeline

if (use_targets) {
  
  df <- targets::tar_read(recoded_data, store = here("_targets"))

} else {
  
  data <- here("data", "processed", "preprocessed_data.csv")
  df <- read_csv(data)
  
}
```

# Investigation over Time

To investigate the changes over time through inferential statistics, logistic regressions were performed for all items. Year was investigated as an ordinal predictor (not a continuous variable), with the year 2022 and 2023 both compared to the baseline from the first wave in 2021. The responses were dichotomized for the outcome into agree/disagree and often/not often (with "sometimes" being coded as "not often"), respectively. Missing and "don't know" responses were filtered out a-priori. McFadden's R^2^ was calculated post-hoc from the model output.

## Open Science

```{r}
os <- c("X38", "X39", "X42", "X43", "X45",
        "X46", "X48", "X49", "X50", "X52")


os_log <- agreement_log_regression(df, os) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")
  
knitr::kable(os_log, format = "html")
```

## Reporting and Academic Integrity

```{r}
raia <- c("X19", "X17", "X18", "X20", "X21",
          "X40", "X44", "X41", "X47")


raia_log <- agreement_log_regression(df, raia) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(raia_log, format = "html")
```

```{r}
raif <- c("X24", "X25", "X26", "X27", "X28", "X29")


raif_log <- frequency_log_regression(df, raif) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(raif_log, format = "html")
```

## Education

```{r}
edua <- c("X16", "X22", "X23", "X51")


edua_log <- agreement_log_regression(df, edua) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(edua_log, format = "html")
```

```{r}
eduf <- c("X30", "X31", "X32")


eduf_log <- frequency_log_regression(df, eduf) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(eduf_log, format = "html")
```

## Institution/Administration

```{r}
adm <- c("X33", "X34", "X35", "X36", "X37")

adm_log <- frequency_log_regression(df, adm) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(adm_log, format = "html")
```

```{r}
hei_log <- agreement_log_regression(df, "X15") %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(hei_log, format = "html")
```

# Models with Academic Role as Additional Predictor

The same models were repeated while adding the academic role in as a predictor. Academic role was dichotomized into staff vs. student (all self-identified students up to ScD level).

## Open Science

```{r}
os_log <- agreement_log_regression_staff(df, os) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")
  
knitr::kable(os_log, format = "html")
```

## Reporting and Academic Integrity

```{r}
raia_log <- agreement_log_regression_staff(df, raia) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(raia_log, format = "html")
```

```{r}
raif_log <- frequency_log_regression_staff(df, raif) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(raif_log, format = "html")
```

## Education

```{r}
edua_log <- agreement_log_regression_staff(df, edua) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(edua_log, format = "html")
```

```{r}
eduf_log <- frequency_log_regression_staff(df, eduf) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(eduf_log, format = "html")
```

## Institution/Administration

```{r}
adm_log <- frequency_log_regression_staff(df, adm) %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(adm_log, format = "html")
```

```{r}
hei_log <- agreement_log_regression_staff(df, "X15") %>% 
  merge(var_overview[, c(1, 2)], ., by.x = "var_id", by.y = "question")

knitr::kable(hei_log, format = "html")
```
