---
title: "Sample Characteristics"
author: "Eva Kormann"
format: html
editor: visual
---

```{r setup, include = FALSE}
library(tidyverse)

# Get data
```

## Number of survey respondents

The survey included three waves, with one wave each year from 2021 to 2023. In addition to a broad national survey, participants were also specifically recruited at the four higher education institutions that are partners within the [OPTIMA](https://lpnu.ua/en/optima) project:

-   Lviv Polytechnic National University (LPU)
-   Sumy State University (SumDU)
-   Vasyl' Stus Donetsk National University (DonNU)
-   Lutsk National Technical University (LutskNTU)

```{r, include = FALSE}
make_table(df$X63, df$X64)
```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
