---
title: "Sample Characteristics"
author:
  - "Eva Kormann"
  - "Thomas Klebel"
format: html
editor: visual
execute: 
  echo: false
---

```{r}
#| output: false
#| warning: false

library(here)
library(tidyverse)
library(janitor)
library(sf)
library(fuzzyjoin)
library(psych)
library(ggridges)
library(patchwork)

options(dplyr.summarise.inform = FALSE)
knitr::opts_chunk$set(dpi = 400)

# Get data

use_targets <- TRUE # Set to false if data is not received by targets pipeline

if (use_targets) {
  
  df <- targets::tar_read(recoded_data, store = here("_targets"))

} else {

  path <- here("data", "processed", "preprocessed_data.csv")
  df <- read_csv(path)
}
```

## Number of survey respondents

The survey included three waves, with one wave each year from 2021 to 2023. Total number of respondents per wave were the following:

-   2021: `r nrow(df[df$X64 == 2021,])` respondents

-   2022: `r nrow(df[df$X64 == 2022,])` respondents

-   2023: `r nrow(df[df$X64 == 2023,])` respondents

In addition to a broad national survey, participants were also specifically recruited at the four higher education institutions that are partners within the [OPTIMA](https://lpnu.ua/en/optima) project:

-   Lviv Polytechnic National University (LPU)
-   Sumy State University (SumDU)
-   Vasyl' Stus Donetsk National University (DonNU)
-   Lutsk National Technical University (LutskNTU)

```{r surveys}

df$X63 <- factor(df$X63, levels = c("National", "LPU", "SumDU", "DonNU", "LutskNTU"))

knitr::kable(tabyl(df, X63, X64) %>% 
               adorn_totals(c("row", "col")),
             format = "html")


ggplot(df, aes(x = X63, fill = as_factor(X64))) +
  geom_bar(position = "dodge", color = "black", linewidth = 0.5) +
  labs(title = "Number of Survey Respondents by Survey and Year",
       x = "Survey", y = "Count", fill = NULL) +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a")) +
  theme(panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"),
        panel.background = element_blank(),
        plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks = element_blank())
```

## Academic role

Participants indicated what they are best described by in terms of academic roles: *"Which of the following best describes you?"*

```{r}
df$X8 <- factor(df$X8, levels = c("student - 1st year", "student - 2nd year", "student - 3rd year", "student - 4th year", "master student", "doctoral student - PhD", "doctoral student - ScD (higher doctorate)", "faculty", "research staff", "university administrator", "librarian", "other"))

roles <- df %>% 
  tabyl(X8, X64) %>% 
  adorn_percentages("col") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(roles, format = "html")
```

## Gender

```{r}
gender <- df %>% 
  tabyl(X9, X64) %>% 
  adorn_percentages("col") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(gender, format = "html")
```

## Degree

*"What is the highest completed academic degree you have?"*

```{r}
df$X10 <- factor(df$X10, levels = c("none (still studying)", "bachelor", "master", "PhD", "ScD (higher doctorate)"))

degrees <- df %>% 
  tabyl(X10, X64) %>% 
  adorn_percentages("col") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(degrees, format = "html")
```

## Work experience

*"How many years have you been working in higher education and/or research?"*

```{r}
df$X11 <- factor(df$X11, levels = c("I have never worked in this sphere", "less than 1 year", "1-5 years", "5-10 years", "more than 10 years"))

experience <- df %>% 
  tabyl(X11, X64) %>% 
  adorn_percentages("col") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(experience, format = "html")
```

## Field/Discipline

*"What is your field of study or work?"*

The table is ordered by the total number of respondents per category across years.

```{r}
field <- as.data.frame.matrix(tabyl(df, X12, X64))
field$X12 <- gsub("\\s*\\(.*?\\)", "", field$X12)

field <- field %>%
  adorn_totals("col") %>% 
  arrange(desc(Total)) %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns() %>% 
  select(-Total)

row.names(field) <- NULL

knitr::kable(field, format = "html")
```

## Publications

*"How many research outputs (articles, monographs, conference papers) have you published in the last 3 years?"*

```{r}
df$X13 <- factor(df$X13, levels = c("0", "1-5", "6-10", "11-20", "21-30", "more than 30"))

publications <- df %>%
  tabyl(X13, X64) %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(publications, format = "html")
```

## Peer review

*"How many peer reviews of research outputs (articles, monographs, reports) have you written in the last 3 years?"*

```{r}
df$X14 <- factor(df$X14, levels = c("0", "1-5", "6-10", "11-20", "21-30", "more than 30"))

review <- df %>%
  tabyl(X14, X64) %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns()

knitr::kable(review, format = "html")
```

## Type of institution

*"What type of higher education institution do you study or work at?"*

The table is ordered by the total number of respondents per category across years.

```{r}
institutions <- df %>%
  tabyl(X54, X64) %>% 
  adorn_totals("col") %>% 
  arrange(desc(Total)) %>% 
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns() %>% 
  select(-Total)

row.names(institutions) <- NULL

knitr::kable(institutions, format = "html")
```

## Sample Comparison over Years

The following figure is intended to explore differences in the sample characteristics between the different years. Multiple characteristics were dichotomized for that purpose (students up to master level vs. others, no degree vs. degree, no peer reviews vs. some, etc.). The figure shows that there was a pronounced increase in 2022 in students and respondents with no degree, work experience, etc. (which also probably relates most to students). The sample in 2023 appears to again be more similar to the 2021 sample.

```{r, sample_years, fig.height = 3.5, fig.width = 5}

years <- df %>% 
  mutate(
    student = case_when(
      str_detect(X8, "student") ~ 1,
      TRUE ~ 0),
    nodegree = case_when(
      X10 == "none (still studying)" ~ 1,
      TRUE ~ 0),
    nowork = case_when(
      X11 == "I have never worked in this sphere" ~ 1,
      TRUE ~ 0),
    nopubs = case_when(
      X13 == "0" ~ 1,
      TRUE ~ 0),
    nopr = case_when(
      X14 == "0" ~ 1,
      TRUE ~ 0)
    ) %>% 
  group_by(X64) %>% 
  summarise_at(c("student", "nodegree", "nowork", "nopubs", "nopr"), mean) %>% 
  pivot_longer(c("student", "nodegree", "nowork", "nopubs", "nopr"),
               names_to = "characteristic")

ggplot(years, aes(x = X64, y = value, color = characteristic)) + 
  geom_line(linewidth = 0.8) +
  geom_point() +
  labs(x = "Survey Year", y = "% respondents", color = "Characteristics") +
  scale_x_continuous(breaks = c(2021, 2022, 2023)) +
  scale_y_continuous(labels = scales::percent, limits = c(0.25, 0.8)) +
  scale_color_brewer(palette = "Dark2",
                     labels = c("No degree", "No peer reviews", "No publications",
                                "No work experience", "Student (up to\nScD level)")) +
  theme(panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"),
        panel.background = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5),
        axis.ticks = element_blank())
```

## Location

*"Your higher education institution is located in"*...

In 2022, this question was extended to *"Your higher education institution is located in (this refers to the original location before 02/24/2022)"*. For the 2023 survey, the phrasing of the question was again reduced to the one first employed in 2021.

Spatial data for the country of Ukraine and its Oblasts were retrieved from [GADM](https://gadm.org/index.html) for visualizations.

```{r}
#| include : false

# Location of institution for mapping
locations <- df %>% 
  group_by(X64, X55) %>% 
  summarize(count = n()) %>% 
  pivot_wider(names_from = X64, values_from = count) %>% 
  rename(NAME_1 = X55) %>% 
  mutate(NAME_1 = gsub(" Oblast", "", NAME_1)) %>% 
  mutate(NAME_1 = gsub("Autonomous Republic of ", "", NAME_1))

# Add Sevastopol for completeness
locations[nrow(locations) + 1,] = list("Sevastopol", 0, 0, 0)

map_data <- here("data", "additional", "map", "gadm41_UKR_1.shp")
ukraine_map <- st_read(map_data, quiet = TRUE) %>% 
  filter(!grepl("\\?", NAME_1)) %>%
  st_make_valid(.)


# Join sample and geographical data by fuzzy string matching
map_counts <- stringdist_join(ukraine_map, locations, 
                              by = "NAME_1",
                              mode = "left",
                              method = "jw",
                              max_dist = 99, 
                              distance_col = "dist") %>%
  group_by(NAME_1.x) %>%
  slice_min(order_by = dist, n = 1)


# Create again shapefile out of merged data
map_counts <- st_as_sf(map_counts)
```

```{r map_2021}
# Plot number of respondents by region in year 2021
map_2021 <- ggplot() +
  geom_sf(data = map_counts, aes(fill = `2021`), color = "white") +
  scale_fill_gradient(low = "white", high = "red", limits = c(0, 2000)) +
  geom_text(data = map_counts,
            aes(label = NAME_1.y,
                x = st_coordinates(st_centroid(geometry))[, "X"],
                y = st_coordinates(st_centroid(geometry))[, "Y"]),
            color = "black", size = 2, check_overlap = FALSE) +
  ggtitle("2021") +
  theme(axis.text = element_blank(),  # Hide axis labels
        axis.ticks = element_blank(), # Hide axis ticks
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 12),
        legend.title = element_blank(), # Hide legend title
        panel.grid = element_blank())  # Hide grid lines
```

```{r map_2022}
# Plot number of respondents by region in year 2022
map_2022 <- ggplot() +
  geom_sf(data = map_counts, aes(fill = `2022`), color = "white") +
  scale_fill_gradient(low = "white", high = "red", limits = c(0, 2000)) +
  geom_text(data = map_counts,
            aes(label = NAME_1.y,
                x = st_coordinates(st_centroid(geometry))[, "X"],
                y = st_coordinates(st_centroid(geometry))[, "Y"]),
            color = "black", size = 2, check_overlap = FALSE) +
  ggtitle("2022") +
  theme(axis.text = element_blank(),  # Hide axis labels
        axis.ticks = element_blank(), # Hide axis ticks
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 12),
        legend.title = element_blank(), # Hide legend title
        panel.grid = element_blank())  # Hide grid lines
```

```{r map_2023}
# Plot number of respondents by region in year 2023
map_2023 <- ggplot() +
  geom_sf(data = map_counts, aes(fill = `2023`), color = "white") +
  scale_fill_gradient(low = "white", high = "red", limits = c(0, 2000)) +
  geom_text(data = map_counts,
            aes(label = NAME_1.y,
                x = st_coordinates(st_centroid(geometry))[, "X"],
                y = st_coordinates(st_centroid(geometry))[, "Y"]),
            color = "black", size = 2, check_overlap = FALSE) +
  ggtitle("2023") +
  theme(axis.text = element_blank(),  # Hide axis labels
        axis.ticks = element_blank(), # Hide axis ticks
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 12),
        legend.title = element_blank(), # Hide legend title
        panel.grid = element_blank())  # Hide grid lines
```

```{r maps_respondents, fig.height = 12, fig.width = 8}
map_2021 / map_2022 / map_2023 +
  plot_layout(guides = "collect")
```

## Displacement

Some universities were also displaced prior to the surveys, some where displaced in between the 2021 and 2022 surveys. In 2022, questions were asked about refugee status, displacement and current whereabouts of the respondent in addition to questions about displacement and status of the respective HEI.

```{r}

df_2022 <- df %>% 
  filter(X64 == "2022")
```

"*Have you been an internally displaced person (within Ukraine) from 24.02.2022 to the present day?*"

```{r}
knitr::kable(df_2022 %>% 
               tabyl(X56),
             format = "html")
```

"*Have you been a refugee abroad from 24.02.2022 to the present day?*"

```{r}
knitr::kable(df_2022 %>% 
               tabyl(X57),
             format = "html")
```

"*Where are you now?*"

```{r}
df_2022$X58 <- factor(
  df_2022$X58, levels = c("at home on the territory controlled by Ukraine",
                     "at home on the temporarily occupied territory",
                     "in another region of Ukraine (internally displaced person)",
                     "abroad as a refugee",
                     "other"))

knitr::kable(df_2022 %>% 
               tabyl(X58),
             format = "html")
```

"*Was your higher education institution in the area of hostilities or under temporary occupation after 02/24/2022?*"

```{r}
#df_2022$X61 <- factor(
#  df_2022$X61, levels = c("no", "yes, now on the territory controlled by Ukraine",
#                     "yes, still on the temporarily occupied territory",
#                     "don’t know"))

knitr::kable(df_2022 %>% 
               tabyl(X61),
             format = "html")
```

"*Is your higher education institution a displaced one?*"

```{r}
#df_2022$X62 <- factor(
#  df_2022$X62, levels = c("no", "yes, after the full-scale invasion in 2022",
#                     "yes, after the start of the war in 2014", "don’t know"))

knitr::kable(df_2022 %>% 
               tabyl(X62),
             format = "html")
```

## Investigation of Response Times

Time stamps were collected within the survey for start and finish. Total response duration was calculated as the difference between those two time stamps in minutes. These were examined more closely to estimate the risk of part of the sample only responding superficially and therefore very quickly.

First, basic descriptive statistics are examined for total response time. Additionally, a simple boxplot is created.

```{r response_boxplot}
knitr::kable(describe(df$duration) %>% 
               mutate_if(is.numeric, ~round(., 2)) %>%
               select(-c(n, trimmed, kurtosis)),
             format = "html")
boxplot(df$duration, main = "Boxplot of Total Response Times", ylab = "Duration")
```

Especially the range of response times from 0 to over 5000 minutes (about three and a half days) is interesting. There are some outliers where respondents might have had the survey open for some days before finally submitting it. The survey took a median time of about 9 minutes to complete.

To have a closer look at the distribution of total response times, a histogram is created. Only times ranging from 0 to 30 minutes will be plotted for better visibility (extreme outliers are omitted).

```{r response_hist}
hist(df$duration[df$duration >= 0 & df$duration <= 30], breaks = 30,
     main = "Histogram of Total Response Times", xlab = "Duration")
```

Only a small share of respondents finished the survey in less then 3 minutes. The peak in the distribution can be seen between 7 and 8 minutes.

Since different groups in the academic context might vary in their familiarity with the topic, they might differ in their response times. To detect possible pronounced differences, distribution of total response time is plotted separately by academic role indicated (again limiting the range of visualized values from 0 to 30).

```{r response_role}
#| warning: false

ggplot(df, aes(x = duration, y = X8)) +
  stat_density_ridges(quantile_lines = TRUE, alpha = 0.7, fill = "lightblue",
                      vline_color = "white") +
  xlim(0, 30) +
  labs(x = "Duration",
       y = "Academic Role",
       title = "Distribution of Total Response Times by Academic Role") +
  scale_y_discrete(labels = c("university administrator" = "administrator",
                              "doctoral student - ScD (higher doctorate)" = "doctoral student - ScD")) +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_line(color = "white")
  )
```

While students have a lower median than, e.g., faculty, there are no major differences in distributions upon closer inspection. The groups do not appear substantially different in their response times.

## Clarity

At the end of the survey, one question asked respondents whether they agreed to following statement: *"All the questions in this questionnaire were clear to me."*

```{r clarity_year}
df$X53 <-  fct_na_value_to_level(df$X53)
knitr::kable(table(df$X53, df$X64), format = "html")


ggplot(df, aes(x = X53, fill = as_factor(X64))) +
  geom_bar(position = "dodge", color = "black", size = 0.5) +
  labs(title = "Clarity of Survey Questions by Year", x = "Clarity",
       y = "Count", fill = NULL) +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a")) +
  theme(panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"),
        panel.background = element_blank(),
        plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5))
```

While the survey questions were clear to the large majority of respondents, there is still a considerable number of people disagreeing on this statement.

```{r clarity_duration}
#| warning: false

ggplot(df, aes(x = duration, y = X53)) +
  stat_density_ridges(quantile_lines = TRUE, alpha = 0.7, fill = "orange",
                      vline_color = "white") +
  xlim(0, 30) +
  labs(x = "Duration",
       y = "Clarity of Questions",
       title = "Distribution of Total Response Times by Clarity Questions") +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_line(color = "white")
  )
```

When investigated in relation to response time, the most notable deviation in distribution can be seen in those not answering the clarity question or answering "don't know" to this question. These respondents tended to finish the survey more quickly.

```{r}

# Calculate respondents indicating low clarity and responding fast
subset_df <- subset(df, duration < 3 & X53 %in% c("strongly disagree", "rather disagree", "NA", "don't know"))

n_fast <- nrow(subset_df)
```

However, the number of respondents who took less than three minutes for the survey and indicated low clarity (or did not answer the clarity question) only lies at `r n_fast`.

## "Don't know" Answers

```{r}

# Calculate number of "don't know" answers
df <- df %>% 
  rowwise() %>% 
  mutate(dontknow = sum(na.omit(c_across(X15:X53)) == "don't know"))


know <- describe(df$dontknow)

corc <- cor(df$duration, df$dontknow, method = "kendall")
```

Participants were able to answer the agreement and frequency questions with "don't know" (39 question in total). Comparing the frequency with which this answer was chosen between groups might also help to investigate response behaviors more closely. Participants responded to a mean of `r round(know$mean, 2)` questions with "don't know" (*SD* = `r round(know$sd, 2)`), ranging from `r know$min` to `r know$max` such answers. Number of these answers is correlated with total response time with a Kendall correlation coefficient of `r round(corc, 2)`. In general, there seems to be no relationship between response time and number of "don't know" responses. On visual inspection, participants that took quite long also responded less with "don't know", however, these participants generally only make up a small part of the sample (median response time is presented in the figure as the red vertical line).

```{r dontknow_duration}
#| warning: false

ggplot(df, aes(x = duration, y = dontknow)) +
  geom_jitter(alpha = .1) +
  geom_smooth(color = "#1065ab") +
  geom_vline(xintercept = know$median, color = "#b31529") +
  xlim(0, 120) +
  labs(x = "Total response time in minutes",
       y = "Number of \"don't know\" responses") +
  theme_minimal()
```

Differences in academic groups might also be of interest here, since some might be more familiar with the topics of the survey than others.

```{r dontknow_role}
#| warning: false

# Plot number of "don't know" answer by academic groups
ggplot(df, aes(x = dontknow, y = X8)) +
  stat_density_ridges(quantile_lines = TRUE, alpha = 0.7, fill = "mediumorchid",
                      vline_color = "white") +
  xlim(0, 30) +
  labs(x = "Number",
       y = "Academic role",
       title = "Distribution of \"don't know\" answers") +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_line(color = "white")
  )
```

In the plot some groups appear to have answered quite rarely with "don't know", mainly university administrators and research staff. Students in their early years of study might also be less familiar with the topic. One notable outlier can be seen in librarians who indicated that they were not able to answer a question the most (however, this is one of the smallest groups among survey respondents).

```{r dontknow_clarity}
#| warning: false

# Plot number of "don't know" answer by response to clarity question
ggplot(df, aes(x = dontknow, y = X53)) +
  stat_density_ridges(quantile_lines = TRUE, alpha = 0.7, fill = "olivedrab3",
                      vline_color = "white") +
  labs(x = "Number",
       y = "Response to clarity question",
       title = "Distribution of \"don't know\" answers") +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_line(color = "white")
  )
```

There seems to be only some difference in the distribution of "don't know" responses in relation to how the clarity questions was answered. Participants that declared that the questions were not clear to them still responded to most of them with something else than "don't know" although to a bit lesser extent. The most notable difference in distribution can be observed for those responding with "don't know" to the clarity question which also responded the same way to almost all content questions. However, this is a comparably small group (see section on clarity question).

## NA answers

```{r}

na <- df %>% 
  summarise(nas = sum(c_across(X38:X52) == "NA"))

df$nas <- na$nas

nas <- describe(df$nas)

corc <- cor(df$duration, df$nas, method = "kendall")
```

Only certain questions could be skipped. Of the content questions, only items X38 to X52 were not mandatory. Therefore a maximum of 15 responses can be missing per person. Participants skipped a mean of `r round(nas$mean, 2)` questions (*SD* = `r round(nas$sd, 2)`), ranging from `r nas$min` to `r nas$max` missings. Number of NAs is correlated with total response time with a Kendall correlation coefficient of `r round(corc, 2)`. In total, missing responses are very rare within the data set.

```{r na_hist}

hist(df$nas, breaks = 15,
     main = "Histogram of NAs", xlab = "Number NAs")
```

## Response Timeline Visualization

```{r response_timeline, fig.height = 2.2, fig.width = 7.5}

ggplot(df, aes(x = X5)) +
  geom_histogram(binwidth = 3600, fill = "black", color = "black") +
  #geom_density(aes(y = ..density..), adjust = .15, fill = "black") +
  geom_vline(xintercept = as.POSIXct(c("2022-01-01", "2023-01-01", "2024-01-01")), 
             linetype = "solid", color = "red") +
  scale_x_datetime(date_labels = "%Y", date_breaks = "1 year",
                   limits = c(as.POSIXct("2021-07-01"),
                              as.POSIXct("2024-03-30"))) +
  labs(x = "Submission of Responses", y = NULL) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
