---
title: "Survey Analysis - plots over time"
author:
  - "Eva Kormann"
  - "Thomas Klebel"
format: html
editor: visual
execute: 
  echo: true
  warning: false
---

```{r}
#| output: false
#| echo: false

library(here)
library(tidyverse)
library(ggchicklet)
library(hrbrthemes)
library(RColorBrewer)

knitr::opts_chunk$set(dpi = 400)

# Get var overview including short titles
vars <- here("data", "additional", "var_short_titles.csv")
var_overview <- read_csv(vars)

# Get functions for analysis
functions <- here("R", "analysis.R")
source(functions)

# Get data
use_targets <- TRUE # Set to false if data is not received by targets pipeline

if (use_targets) {
  
  df <- targets::tar_read(recoded_data, store = here("_targets"))

} else {
  
  additional_functions <- here("R", "functions.R")
  source(additional_functions)
  
  data <- here("data", "processed", "preprocessed_data.csv")
  df <- read_csv(data) %>% 
    likert_to_factor(., var_overview)
  
}
```

```{css, echo = FALSE}
/* Make tables scroll */ 
  
  .scroll {
    max-height: 500px;
    overflow-y: scroll;
  }
```

```{r, echo=FALSE}
# Define item groups
os <- c("X38", "X39", "X42", "X43", "X45",
        "X46", "X48", "X49", "X50", "X52")

raia <- c("X19", "X17", "X18", "X20", "X21",
          "X40", "X44", "X41", "X47")

raif <- c("X24", "X25", "X26", "X27", "X28", "X29")

edua <- c("X16", "X22", "X23", "X51")

eduf <- c("X30", "X31", "X32")

adm <- c("X33", "X34", "X35", "X36", "X37")
```

# Development over Time

Since the survey was conducted in three separate waves (2021, 2022, 2023), responses are also compared between those years. Similar to the overview plots where "don't know" and missing responses were separated out before calculating proportions of certain response categories, these are also filtered out here. Percentages therefore refer to all respondents that chose one of the response categories presented in the figure.

## Research

### Open Science

```{r os_development, fig.height = 10, fig.width = 7.5}
out <- plot_time(df, var_overview, os, type = "agreement")

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```

### Reporting and Academic Integrity

```{r raia_development, fig.height = 7.8, fig.width = 7.5}
out <- plot_time(df, var_overview, raia, type = "agreement")

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```

```{r raif_development, fig.height = 5.8, fig.width = 7.5}
out <- plot_time(df, var_overview, raif, type = "frequency")

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```

## Education

```{r edua_development, fig.height = 5.5, fig.width = 5.5}
out <- plot_time(df, var_overview, edua, type = "agreement", ncol = 2,
                 var_wrap = 40)

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```

```{r eduf_development, fig.height = 3, fig.width = 7.5}
out <- plot_time(df, var_overview, eduf, type = "frequency")

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```

## Institution/Administration

```{r adm_development, fig.height = 5, fig.width = 7.5}
out <- plot_time(df, var_overview, adm, type = "frequency")

out$p
```

```{r}
#| class: scroll

knitr::kable(out$data)
```
