---
title: "Survey Analysis"
author:
  - "Eva Kormann"
  - "Thomas Klebel"
format: html
editor: visual
execute: 
  echo: true
  warning: false
---

```{r}
#| output: false
#| echo: false

library(here)
library(tidyverse)
library(psych)
library(ggchicklet)
library(patchwork)
library(janitor)
library(hrbrthemes)
library(RColorBrewer)
library(broom)

knitr::opts_chunk$set(dpi = 400)

# Get var overview including short titles
vars <- here("data", "additional", "var_short_titles.csv")
var_overview <- read_csv(vars)

# Get functions for analysis
functions <- here("R", "analysis.R")
source(functions)

# Get data
use_targets <- TRUE # Set to false if data is not received by targets pipeline

if (use_targets) {
  
  df <- targets::tar_read(recoded_data)

} else {
  
  data <- here("data", "processed", "preprocessed_data.csv")
  df <- read_csv(data)
  
}
```

# Overview

The following plots give a top-level overview over the responses to all questions, grouped by topics. They are also differentiated between questions that were about the extent of agreement with certain statements and questions that asked about the estimated frequency with which certain practices occur at the institution. The survey was conducted in three waves in three subsequent years, but all responses are combined for the overview plots.

## Research

### Open Science

Multiple questions within the survey inquired about how much participants agreed with certain statements related to different Open Science practices. The full statements in this group are presented in the table below.

```{r}
os <- c("X38", "X39", "X42", "X43", "X45",
        "X46", "X48", "X49", "X50", "X52")

knitr::kable(var_overview[var_overview$var_id %in% os, c(1, 2, 4)])
```

In the following figure, responses are visualized to the Open Science questions. They are sorted by the amount of agreement with these statements. Proportions of the different answer categories were calculated after removing missing and "don't know" responses. These responses are plotted separately on the right hand side, indicating their percentage of the total responses.

```{r os_overview, fig.height = 7.5, fig.width = 9}
plot_agreement_overview(df, var_overview, os)
```

### Reporting and Academic Integrity

Nine items inquire about the agreement with statements related to reporting and academic integrity. The full items are presented in the table below.

```{r}
raia <- c("X19", "X17", "X18", "X20", "X21",
          "X40", "X44", "X41", "X47")
knitr::kable(var_overview[var_overview$var_id %in% raia, c(1, 2, 4)])
```

Caution: Question X19 ("Changes to the experimental data \[...\] are acceptable") is the only statement phrased in the opposite direction. This needs to be considered when interpreting the responses.

```{r raia_overview, fig.height = 7, fig.width = 9}
plot_agreement_overview(df, var_overview, raia)
```

Six questions were about the frequency with which certain questionable research practices around reporting occur at the respondent's institution (see table below).

```{r}
raif <- c("X24", "X25", "X26", "X27", "X28", "X29")
knitr::kable(var_overview[var_overview$var_id %in% raif, c(1, 2, 4)])
```

The following overview plot is sorted by the proportion of "never" and "rarely" responses. These questions could not be skipped, therefore there are no missing answers.

```{r raif_overview, fig.height = 5, fig.width = 9}
plot_frequency_overview(df, var_overview, raif)
```

## Education

While most questions referred to research, others focused more on education at the respective HEI. Respondents were inquired about their agreement with certain education-related statements (see table below), most of them also related to Open Science, but in an educational context.

```{r}
edua <- c("X16", "X22", "X23", "X51")
knitr::kable(var_overview[var_overview$var_id %in% edua, c(1, 2, 4)])
```

```{r edua_overview, fig.height = 4, fig.width = 9}
plot_agreement_overview(df, var_overview, edua, sort = FALSE)
```

The occurrence of three different types of student misconduct were also investigated within the survey (see table and figure below).

```{r}
eduf <- c("X30", "X31", "X32")
knitr::kable(var_overview[var_overview$var_id %in% eduf, c(1, 2, 4)])
```

```{r eduf_overview, fig.height = 3.5, fig.width = 9}
plot_frequency_overview(df, var_overview, eduf)
```

## Institution/Administration

The survey contained five questions on how misconduct is handled at the respondent's respective institution. Respondents indicated the frequency with which certain practices occurred at their institution.

```{r}
adm <- c("X33", "X34", "X35", "X36", "X37")
knitr::kable(var_overview[var_overview$var_id %in% adm, c(1, 2, 4)])
```

```{r adm_overview, fig.height = 4.5, fig.width = 9}
plot_frequency_overview(df, var_overview, adm)
```

One general item also inquired about the following statement: "Overall, the processes at my higher education institution are organised in a clear and transparent manner". The agreement is visualized by the different institutional vs. the national survey.

```{r hei_processes, fig.height = 4.5, fig.width = 9}
df$X63 <- factor(df$X63,
                 levels = c("National", "LPU", "SumDU", "DonNU", "LutskNTU"))
plot_agreement(df, var_overview, "X15", group = "X63", xlim = 3)
```

# Development over Time

Since the survey was conducted in three separate waves (2021, 2022, 2023), responses are also compared between those years. Similar to the overview plots where "don't know" and missing responses were separated out before calculating proportions of certain response categories, these are also filtered out here. Percentages therefore refer to all respondents that chose one of the response categories presented in the figure.

## Research

### Open Science

```{r os_development, fig.height = 15, fig.width = 8}

plots <- list()

for (q in os) {
  if (q %in% paste0("X", 38:53)) {
    legend <- TRUE
  } else {legend <- FALSE}
  plot <- plot_agreement_line(df, question = q, legend)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

### Reporting and Academic Integrity

```{r raia_development, fig.height = 15, fig.width = 8}
 
plots <- list()

for (q in raia) {
  if (q %in% paste0("X", 38:53)) {
    legend <- TRUE
  } else {legend <- FALSE}
  plot <- plot_agreement_line(df, question = q, legend)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

```{r raif_development, fig.height = 9, fig.width = 8}

plots <- list()

for (q in raif) {
  plot <- plot_frequency_line(df, question = q, legend = TRUE, ylim = 50)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

### Education

```{r edua_development, fig.height = 5.5, fig.width = 8}
 
plots <- list()

for (q in edua) {
  if (q %in% paste0("X", 38:53)) {
    legend <- TRUE
  } else {legend <- FALSE}
  plot <- plot_agreement_line(df, question = q, legend)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

```{r eduf_development, fig.height = 5.5, fig.width = 8}

plots <- list()

for (q in eduf) {
  plot <- plot_frequency_line(df, question = q, legend = TRUE, ylim = 40)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

### Institution/Administration

```{r adm_development, fig.height = 9, fig.width = 8}

plots <- list()

for (q in adm) {
  plot <- plot_frequency_line(df, question = q, legend = TRUE)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10))
```

An alternative plot type to the line plot is the area plot. For the following exemplary item (X15: "Overall, the processes at my higher education institution are organised in a clear and transparent manner") both plots are shown for comparison.

```{r hei_processes_comparison, fig.height = 3.2, fig.width = 8}

line <- plot_agreement_line(df, "X15", FALSE)
area <- plot_agreement_area(df, "X15", TRUE)

line + area
```

# Comparison Between Groups

To compare agreement over time across groups, the responses were dichotomized into agreement ("rather agree", "strongly agree") and disagreement ("rather disagree", "strongly disagree"). This way the percentage of respondents agreeing with a certain statement can be compared across groups and over time. Frequency responses were recoded into "often" ("very often", "frequently") and "not often" ("sometimes", "rarely", "never"). Missing responses and "don't know" answers were removed before the percentages were calculated. As in the previous section, percentages refer to all respondents that chose one of the response categories presented (and not "don't know").

## Academic Roles

To compare respondents with different academic roles, students are first summarized for purposes of visualization (see below for differentiated visualizations for this group). The "other" category was additionally excluded, since it is quite small and difficult to characterize. Some groups differ quite substantially in their size (e.g., lot of faculty, only few librarians). This has to be taken into account for interpretation.

```{r}

roles <- df %>% 
  filter(X8 != "other") %>% 
  mutate(roles = case_when(
    str_detect(X8, "student.*year") ~ "student",
    X8 == "master student" ~ "student",
    X8 == "doctoral student - ScD (higher doctorate)" ~ "doctoral student - ScD",
    TRUE ~ X8),
  roles = factor(roles, levels = c("student", "doctoral student - PhD",
                                   "doctoral student - ScD",
                                   "faculty", "research staff",
                                   "university administrator", "librarian")))
```

### Open Science

```{r os_roles, fig.height = 18, fig.width = 8}

plots <- list()

for (q in os) {
  plot <- plot_agreement_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Reporting and Academic Integrity

```{r raia_roles, fig.height = 18, fig.width = 8}

plots <- list()

for (q in raia) {
  plot <- plot_agreement_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r raif_roles, fig.height = 10, fig.width = 8}

plots <- list()

for (q in raif) {
  plot <- plot_frequency_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Education

```{r edua_roles, fig.height = 7, fig.width = 8}

plots <- list()

for (q in edua) {
  plot <- plot_agreement_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r eduf_roles, fig.height = 7, fig.width = 8}

plots <- list()

for (q in eduf) {
  plot <- plot_frequency_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Institution/Administration

```{r, adm_roles, fig.height = 10, fig.width = 8}

plots <- list()

for (q in adm) {
  plot <- plot_frequency_groups(roles, question = q, "roles")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r, hei_processes_roles, fig.height = 3.5, fig.width = 5}

plot_agreement_groups(roles, "X15", "roles")
```

## Students by Year

In the plots above, students were all grouped into one category. Since it is of interest how students' perceptions differ depending on the stage of their studies, they are compared more closely in the subsequent figures (these could maybe added to the appendix).

```{r}

students <- df %>% 
  filter(str_detect(X8, "student.*year") | X8 == "master student")
```

### Open Science

```{r os_students, fig.height = 18, fig.width = 8}

plots <- list()

for (q in os) {
  plot <- plot_agreement_groups(students, question = q, "X8", TRUE, TRUE)
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Reporting and Academic Integrity

```{r raia_students, fig.height = 18, fig.width = 8}

plots <- list()

for (q in raia) {
  plot <- plot_agreement_groups(students, question = q, "X8")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r raif_students, fig.height = 10, fig.width = 8}

plots <- list()

for (q in raif) {
  plot <- plot_frequency_groups(students, question = q, "X8")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Education

```{r edua_students, fig.height = 7, fig.width = 8}

plots <- list()

for (q in edua) {
  plot <- plot_agreement_groups(students, question = q, "X8")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r eduf_students, fig.height = 7, fig.width = 8}

plots <- list()

for (q in eduf) {
  plot <- plot_frequency_groups(students, question = q, "X8")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

### Institution/Administration

```{r, adm_students, fig.height = 10, fig.width = 8}

plots <- list()

for (q in adm) {
  plot <- plot_frequency_groups(students, question = q, "X8")
  plots[[q]] <- plot
}

wrap_plots(plots, ncol = 2) +
  plot_layout(guides = "collect") & 
  theme(plot.tag = element_text(size = 10))
```

```{r, hei_processes_students, fig.height = 3.5, fig.width = 5}

plot_agreement_groups(students, "X15", "X8")
```

# Investigation through Inferential Statistics

To investigate the changes over time through inferential statistics, logistic regressions were performed for all items. Year was investigated as an ordinal predictor (not a continuous variable), with the year 2022 and 2023 both compared to the baseline from the first wave in 2021. The responses were dichotomized for the outcome into agree/disagree and often/not often (with "sometimes" being coded as "not often"), respectively. Missing and "don't know" responses were filtered out a-priori. McFadden's R^2^ was calculated post-hoc from the model output.

## Open Science

```{r}
os_log <- agreement_log_regression(df, os) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")
  
knitr::kable(os_log, format = "html")
```

## Reporting and Academic Integrity

```{r}
raia_log <- agreement_log_regression(df, raia) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(raia_log, format = "html")
```

```{r}
raif_log <- frequency_log_regression(df, raif) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(raif_log, format = "html")
```

## Education

```{r}
edua_log <- agreement_log_regression(df, edua) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(edua_log, format = "html")
```

```{r}
eduf_log <- frequency_log_regression(df, eduf) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(eduf_log, format = "html")
```

## Institution/Administration

```{r}
adm_log <- frequency_log_regression(df, adm) %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(adm_log, format = "html")
```

```{r}
hei_log <- agreement_log_regression(df, "X15") %>% 
  merge(var_overview[, c(1,4)], ., by.x = "var_id", by.y = "question")

knitr::kable(hei_log, format = "html")
```
